---
title: "Final Project Report"
output: github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Data Preparation and Exploration
0.0, Load Packages
```{r load packages and function}
library(data.table)
library(DT)
library(dplyr)
library(ggplot2)
```
0.1, Read Dataset
```{r read dataset}
#read train, test and submission dataset
train_dt <- fread(input = '../R data project/train.csv')
test_dt <- fread(input = '../R data project/test.csv')
sample_submission_dt <- fread(input = '../R data project/sample_submission.csv')

#read relevant data table
holidays_events_dt <- fread(input = '../R data project/holidays_events.csv')
items_dt <- fread(input = '../R data project/items.csv')
oil_dt <- fread(input = '../R data project/oil.csv')
stores_dt <- fread(input = '../R data project/stores.csv')
transactions_dt <- fread(input = '../R data project/transactions.csv')
```

0.2, Sample train dataset
```{r sample train data}
sample.size <- 100000
sample_train_dt <- sample_n(train_dt, sample.size)
```
0.3, Help function
```{r help function}
plot.categoric <- function(dt,col,color = 2){
  sample <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
  num.plot <- ggplot(dt, aes(x=get(col))) + 
    geom_bar(fill = sample[color], alpha = 0.7) + 
    geom_text(aes(label = ..count..), stat='count', vjust=-0.5) + 
    theme_minimal() + 
    xlab(col) +
    ylab("Freq") +
    theme(axis.text.x = element_text(angle = 15, size=10))
  return(num.plot)
}

plot.histogram <- function(dt,col,color = 5){
  sample <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
  num.plot <- ggplot(dt, aes(x=get(col))) + 
    geom_histogram(breaks=seq(0,70,by=10), fill = sample[color], alpha = 0.8) + 
    theme_minimal() + 
    xlab(col)
    theme(axis.text.x = element_text(angle = 15, size=10))
  return(num.plot)
}

plot.scatter <- function(dt,x,y,color = 3){
  sample <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
  num.plot <- ggplot(dt, aes(x=get(x),y=get(y))) + 
    geom_point(size = 1, color = sample[color] ) +
    geom_line() +
    xlab(x)+
    ylab(y)
  return(num.plot)
}

#define color
color <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

1, Variable Identification
```{r check train dataset summary}
summary(sample_train_dt)
min.train.id <- min(sample_train_dt$id)
max.train.id <- max(sample_train_dt$id)
min.date <- min(sample_train_dt$date)
max.date <- max(sample_train_dt$date)
```
```{r check train dataset}
first.lines <- 20
datatable(sample_train_dt[1:first.lines,])
```
## About the data
**id**: Unique identification for records, train dataset ranges from `r min.train.id` to `r max.train.id`.
**date**: Date information of the records, train dataset ranges from `r min.date` to `r max.date`.
**store_nbr**: Unique identification for store number.
**item_nbr**: Unique identification for item number.
**unit_sales**: The units of sales by every record. (Removing negative values)
**onpromotion**: Catehorical variable indicates that items sold on protion or not.
```{r explore store number}
store_nbr.plot <- ggplot(sample_train_dt, aes(x=store_nbr)) +
  geom_bar(aes(weight=rep(1:length(store_nbr)),fill= color[1])) +
  scale_y_continuous(labels = scales::unit_format()) +
  xlab("Store Number") +
  ylab("Sales") +
  ggtitle("Sales in Distinct Store")
store_nbr.plot
```
```{r explore item number}
top.item.sales <- {sample_train_dt[, .N, by=c("item_nbr")] %>% data.table()} [order(-N)]
datatable(top.item.sales[1:20,])
```
```{r explore sales}
promotion.plot <- ggplot(sample_train_dt, aes(x=unit_sales)) + 
    geom_histogram(breaks=seq(0,20,by=2), fill = color[5], alpha = 0.7) + 
    theme_minimal() + 
    xlab("positive unit_sales")
promotion.plot
```

```{r explore promotion}
sale.plot <- ggplot(sample_train_dt, aes(x=onpromotion)) + 
    geom_bar(fill = color[2], alpha = 0.5, width = 0.5) + 
    geom_text(aes(label = ..count..), stat='count', vjust=-0.5) + 
    theme_minimal() + 
    xlab("ompromotion") +
    ylab("Freq")+
    theme(axis.text.x = element_text(angle = 15, size=10))
sale.plot
```

```{r check test dataset}
summary(test_dt)
test.row <- nrow(test_dt)
```
In total `r test.row` lines of test data, having identical table structure with train dataset except for lacking *unit_sales* variable.

```{r check holidays}
datatable(holidays_events_dt)
```
**date**: Holiday date information of the records.
**type**: Indication of the holiday's type.
**locale**: Indication of holiday locality attributes.
**transferred**: Whether the holiday is transfered.
```{r explore holiday type}
holiday.type.plot <- ggplot(holidays_events_dt, aes(x=type)) + 
    geom_bar(fill = color[3], alpha = 0.7, width = 0.5) + 
    geom_text(aes(label = ..count..), stat='count', vjust=-0.5) + 
    theme_minimal() + 
    xlab("Type") +
    ylab("Freq")+
    theme(axis.text.x = element_text(angle = 15, size=10))
holiday.type.plot
```
```{r explore locale}
holiday.locale.plot <- ggplot(holidays_events_dt, aes(x=locale)) + 
    geom_bar(fill = color[4], alpha = 0.8, width = 0.5) + 
    geom_text(aes(label = ..count..), stat='count', vjust=-0.5) + 
    theme_minimal() + 
    xlab("Locale") +
    ylab("Freq")+
    theme(axis.text.x = element_text(angle = 15, size=10))
holiday.locale.plot
```
```{r explore holiday tranfer}
holiday.transferred.plot <- ggplot(holidays_events_dt, aes(x=transferred)) + 
    geom_bar(fill = color[2], alpha = 0.7, width = 0.5) + 
    geom_text(aes(label = ..count..), stat='count', vjust=-0.5) + 
    theme_minimal() + 
    xlab("Transferred") +
    ylab("Freq")+
    theme(axis.text.x = element_text(angle = 15, size=10))
holiday.transferred.plot
```


```{r check items}
datatable(items_dt[1:first.lines,])
```
**item_nbr**: The item number.
**family**: The category of item according to number.
**class**: The class of item.
**perishable**: Whether the item is perishable.
```{r explore item family}
largest.item.family.plot <- ggplot(items_dt, aes(x=family)) + 
    geom_bar(fill = color[4], alpha = 0.8, width = 0.5) + 
    geom_text(aes(label = ..count..), stat='count', vjust=-0.5) + 
    theme_minimal() + 
    xlab("Family") +
    ylab("Freq")+
    theme(axis.text.x = element_text(angle = 15, size=10))
largest.item.family.plot
```
```{r explore item parishable}
item.perishable.plot <- ggplot(items_dt, aes(x=perishable)) + 
    geom_bar(fill = color[2], alpha = 0.7, width = 0.5) + 
    geom_text(aes(label = ..count..), stat='count', vjust=-0.5) + 
    theme_minimal() + 
    xlab("Perishable") +
    ylab("Freq")+
    theme(axis.text.x = element_text(angle = 15, size=10))
item.perishable.plot 
```

```{r check oil price}
price.plot <- ggplot(oil_dt, aes(x=date,y=dcoilwtico)) + 
    geom_point(size = 1, color = color[3] ) +
    geom_line() +
    xlab("date")+
    ylab("dcoilwtico")
price.plot
```
```{r check stores}
datatable(stores_dt)
```
```{r transactions}
transactions_dt[,lapply(.SD, FUN="sum"), .SDcols=c("transactions"), by=c("store_nbr")]

```


